---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CCDB

<!-- badges: start -->
<!-- badges: end -->

The goal of CCDB is to not only make it easier for a user to acquire data from the consumer complaints database, it also provides useful information and summaries. It entails multiple functions that tailors its search, and produces an output of either a dataframe or a visualization chart/graph depending on the function. Even though all functions are made to work on their own, some functions are better used in combination with others.

# Vignette
Please refer to the vignette for a more complete and detailed information about the original consumer complaints database, rationale for the API R package, individual functions, and examples.


## Examples

```{r example}
library(CCDB)
# example of product function for debt collection
product_function(product="Debt collection",subproduct_option = TRUE,state_option=TRUE)

#example of general function for Bank of America, National Association
general_function(company = "BANK OF AMERICA, NATIONAL ASSOCIATION",product = "Checking or savings account",issue="Managing an account", limit=10)

#example of visualization tool using sample df
visualization_tool_function()

#example of visualization tool using the outputs of product function for Debt collection
example1<- product_function(product="Debt collection")
#using the subproduct category dataframe from the output
visualization_tool_function(example1[1],limit=10)

#example of heatmap state visualization tool using sample df
heatmap_state_function()

#using the output of product function for Student loan
example2 <- product_function(product="Vehicle loan or lease")

#selecting the second dataframe, as that is the one with the state attributes and values
heatmap_state_function(example2[2])
```
